---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

```{r}

library(tidyverse)
library(ggpubr)


dat <- read.csv("activity.csv")
```


## What is mean total number of steps taken per day?

```{r}


datgrouped <- dat %>%
  group_by(date)


  summ <- datgrouped %>% summarise(
    summ = sum(steps)
  )
  
  colnames(summ) <- c("date", "total number steps")
  
  plot1 <- ggplot(summ, aes(`total number steps`)) + 
  geom_histogram(binwidth = 500) 
  plot1
  
 avg <- as.integer(mean(summ$`total number steps`, na.rm = TRUE))
 me <- median(summ$`total number steps`, na.rm = TRUE)
  
```
The mean total number of steps taken per day is: `r avg` and the median total number of steps taken per day is: `r me`

## What is the average daily activity pattern?

```{r}
dat2 <- dat

datgrouped2 <- dat2 %>%
  group_by(interval)


  avg2 <- datgrouped2 %>% summarise(
    mean_steps = mean(steps, na.rm = TRUE)
  )

p <- ggplot(avg2, aes(x=interval, y=mean_steps)) +
        geom_line(color="#69b3a2") + 
        xlab("") +
        ylab("Average number of steps per invervall")
        p

max <- avg2 %>% filter(mean_steps == max(mean_steps))
max1 <- max[[1]]


```

The 5 minute interval with the the most steps on average across all the days in the data set is: `r max1`

## Imputing missing values

```{r}
#total number of missing values

mis <- is.na(dat$steps)
mis2 <- sum(mis)

#impute values

dat3 <- merge(avg2, dat, )

dat3 <- merge(dat, avg2, 
                  by.x = "interval", 
                  by.y = "interval", all = FALSE)

#new data set with imputed values
dat3$steps <- ifelse(is.na(dat3$steps), dat3$mean_steps, dat3$steps)


#plot 
dat3$date <- as.Date(dat3$date)

datgroupedIM <- dat3 %>%
  group_by(date)


  summIM <- datgroupedIM %>% summarise(
    summIM = sum(steps)
  )
  
  colnames(summIM) <- c("date", "total number steps")
  
  plot3 <- ggplot(summIM, aes(`total number steps`)) + 
  geom_histogram(binwidth = 500) 
  plot3
  
 avgImp <- as.integer(mean(summIM$`total number steps`))
 meImp <- as.integer(median(summIM$`total number steps`))


```

The total number of missing values in the data set is: `r mis2`

The strategy to impute missing values will be to replace them with the global average of the same interval.

After imputing the mean total number of steps taken per day is: `r avgImp` and the median total number of steps taken per day is: `r meImp`. The mean values as unaffected, the median value changed from `r me` to `r meImp`



## Are there differences in activity patterns between weekdays and weekends?


```{r}

dat4 <- dat3

dat4 <- dat4 %>% 
        mutate(date = weekdays(date))

Weekend <- filter(dat4, date == "Sunday" | date == "Saturday")
Workdays <- filter(dat4, date == "Monday" | date == "Tuesday" | date == "Wednesday"| date == "Thursday" | date == "Friday")

Wendgrouped <- Weekend %>%
  group_by(interval)

Wendavg <- Wendgrouped %>% summarise(
    mean_steps = mean(steps, na.rm = TRUE)
  )


Workgrouped <- Workdays %>%
  group_by(interval)

Workavg <- Workgrouped %>% summarise(
    mean_steps = mean(steps, na.rm = TRUE)
  )


plot4 <- ggplot(Wendavg, aes(x=interval, y=mean_steps)) +
        geom_line(color="#69b3a2") + 
        xlab("") +
        ylab("Number of steps") +
        labs(title = "Weekend")
        


plot5 <- ggplot(Workavg, aes(x=interval, y=mean_steps)) +
        geom_line(color="#69b3a2") + 
        xlab("") +
        ylab("Number of steps") +
        labs(title = "Week days")




figure <- ggarrange(plot4, plot5,
                    labels = c("A", "B"),
                    ncol = 1, nrow = 2)
figure


```




